---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# statBasics

`R` package of the course **Métodos Estatísticos** at Federal University of Bahia.

## Confidence Interval

Here, we compute the confidence interval using the methods as presented by Montgomery and Runger (2010). All methods implemented are slighted modification of methods already implemented in `stats`. The user can compute bilateral and unilateral confidence intervals.

### Confidence Interval for Proportion

There are three approaches to compute confidence interval for proportion in this package. 

#### Number of success in `n` (scalar value) trials

```{r}
library(tidyverse)
library(statBasics)
size  <- 1000
sample <- rbinom(size, 1, prob = 0.5)
n_success <- sum(sample)
ci_bern(n_success, size, conf_level = 0.99)
```

#### Number of success in a vector

```{r}
library(tidyverse)
library(statBasics)
n <- c(30, 20, 10)
x <- n %>% map_int(~ sum(rbinom(1, size = .x, prob = 0.75)))
ci_bern(x, n, conf_level = 0.99)
```

#### Vector of success

```{r}
library(tidyverse)
library(statBasics)
x <- rbinom(50, size = 1, prob = 0.75)
ci_bern(x, conf_level = 0.99)
```

### Confidence interval of mean (normal Distribution)

Let's illustrated with a simulated cases the confidence interval for two cases: 1)  standard deviation is known; 2) standard deviation is not known.

#### Standard Deviation is unknown

```{r}
library(tidyverse)
library(statBasics)
media_pop <- 10
sd_pop <- 2
x <- rnorm(100, mean = media_pop)
ci_norm(x, conf_level = 0.91)
```

#### Standard Deviation is known

```{r}
library(tidyverse)
library(statBasics)
media_pop <- 10
sd_pop <- 2
x <- rnorm(100, mean = media_pop, sd = sd_pop)
ci_norm(x, sd_pop = sd_pop, conf_level = 0.91)
```

### Confidence interval for standard deviation (normal distribution)

```{r}
library(tidyverse)
library(statBasics)
media_pop <- 10
sd_pop <- 2
x <- rnorm(100, mean = media_pop, sd = sd_pop)
ci_norm(x, parameter = 'variance', conf_level = 0.91)
```

### Confidence interval for mean (exponential distribution)

```{r}
library(tidyverse)
library(statBasics)
media_pop <- 800
taxa_pop <- 1 / media_pop
x <- rexp(100, rate = taxa_pop)
ci_exp(x)
```

### Confidence interval for mean

Confidence interval using *t* distribution is suitable, even the distribution is not normal, as illustrated in the example bellow.

```{r}
library(tidyverse)
library(statBasics)
media_pop <- 50
x <- rpois(100, lambda  = media_pop)
ci_general(x)
```

## Hypothesis testing

In the following, we will illustrate the use of this package to test scientific hypothesis for one sample with examples. All methods are already implemented in `R`, I have only made slightly modifications to teaching purposes.

### Hypothesis testing for mean

In the examples below, `mean_null` is the mean in the null hypothesis `H0`:

1. `alternative == "two.sided"`: `H0: mu == mean_null` and `H1: mu != mean_null`. Default value.
2. `alternative == "less"`: `H0: mu >= mean_null` and `H1: mu < mean_null`
3. `alternative == "greater"`: `H0: mu =< mean_null` and `H1: mu > mean_null`

#### Normal distribution with known variance

```{r}
library(tidyverse)
library(statBasics)
mean_null <- 5
sd_pop <- 2
x <- rnorm(100, mean = 10, sd = sd_pop)
ht_1pop_mean(x, mu = mean_null, conf_level = 0.95, sd_pop = sd_pop, alternative = "two.sided")
```

#### Normal distribution with unknown variance

```{r}
library(tidyverse)
library(statBasics)
mean_null <- 5
sd_pop <- 2
x <- rnorm(100, mean = 10, sd = sd_pop)
ht_1pop_mean(x, mu = mean_null, conf_level = 0.95, sd_pop = sd_pop, alternative = "two.sided")
```

### Hypothesis testing for variance

In the examples below, `sigma_null` is the standard deviation in the null hypothesis `H0`:

1. `alternative == "two.sided"`: `H0: sigma == sigma_null` and `H1: sigma != sigma_null`. Default value.
2. `alternative == "less"`: `H0: sigma >= sigma_null` and `H1: sigma < sigma_null`
3. `alternative == "greater"`: `H0: sigma =< sigma_null` and `H1: sigma > sigma_null`

```{r}
library(tidyverse)
library(statBasics)
sigma_null <- 4
sd_pop <- 2
x <- rnorm(100, mean = 10, sd = sd_pop)
ht_1pop_var(x, sigma = sigma_null, conf_level = 0.95, alternative = "two.sided")

```

### Hypothesis testing for proportion

In the examples below, `proportion_null` is the standard deviation in the null hypothesis `H0`:

1. `alternative == "two.sided"`: `H0: proportion == proportion_null` and `H1: proportion != proportion_null`. Default value.
2. `alternative == "less"`: `H0: proportion >= proportion_null` and `H1: proportion < proportion_null`
3. `alternative == "greater"`: `H0: proportion =< proportion_null` and `H1: proportion > proportion_null`

#### Number of success

You can have the number of success (in a number of trials) as a scalar.

```{r}
library(tidyverse)
library(statBasics)
proportion_null <- 0.1
p0 <- 0.75
x <- rbinom(1, size = 1000, prob = p0)
ht_1pop_prop(x, 1000, proportion = p0, alternative = "two.sided", conf_level = 0.95)
```

#### Number of success in a vector

You can have the number of success as a vector. You must also give the vector of number of trials.

```{r}
library(tidyverse)
library(statBasics)
proportion_null <- 0.9
p0 <- 0.75
n <- c(10, 20, 30)
x <- n %>% map_int(~ rbinom(1, .x, prob = p0))
ht_1pop_prop(x, n, proportion = p0, alternative = "less", conf_level = 0.99)
```

#### Vector of success (0 or 1)

You can have the number of success (zero ou one).

```{r}
library(tidyverse)
library(statBasics)
proportion_null <- 0.1
p0 <- 0.75
x <- rbinom(1000, 1, prob = p0)
ht_1pop_prop(x, proportion = p0, alternative = "greater", conf_level = 0.95)
```
