---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# statBasics

`R` package of the course **Métodos Estatísticos** at Federal University of Bahia.

# Confidence Interval

Here, we compute the confidence interval using the methods as presented by Montgomery and Runger (2010). All methods implemented are slighted modification of methods already implemented in `stats`. The user can compute bilateral and unilateral confidence intervals.

## Confidence Intervalo for Proportion

There are three approaches to compute confidence interval for proportion in this package. 

### Number of sucess in `n` (scalar value) trials

```{r}
library(tidyverse)
library(statBasics)
size  <- 1000
sample <- rbinom(size, 1, prob = 0.5)
n_success <- sum(sample)
ci_bern(n_success, size, conf_level = 0.99)
```

### Number of sucess in a vector

```{r}
library(tidyverse)
library(statBasics)
n <- c(30, 20, 10)
x <- n %>% map_int(~ sum(rbinom(1, size = n[.x], prob = 0.75)))
ci_bern(x, n, conf_level = 0.99)
```

### Vector of success

```{r}
library(tidyverse)
library(statBasics)
x <- rbinom(50, size = 1, prob = 0.75)
ci_bern(x, conf_level = 0.99)
```

## Confidence interval of mean (normal Distribution)

Let's illustrated with a simulated cases the confidence interval for two cases: 1)  standard deviation is known; 2) standard deviation is not known.

### Standard Deviation is unknown

```{r}
library(tidyverse)
library(statBasics)
media_pop <- 10
sd_pop <- 2
x <- rnorm(100, mean = media_pop)
ci_norm(x, conf_level = 0.91)
```

### Standard Deviation is known

```{r}
library(tidyverse)
library(statBasics)
media_pop <- 10
sd_pop <- 2
x <- rnorm(100, mean = media_pop, sd = sd_pop)
ci_norm(x, sd_pop = sd_pop, conf_level = 0.91)
```

## Confidence interval for standard deviation (normal distribution)

```{r}
library(tidyverse)
library(statBasics)
media_pop <- 10
sd_pop <- 2
x <- rnorm(100, mean = media_pop, sd = sd_pop)
ci_norm(x, parameter = 'variance', conf_level = 0.91)
```

## Confidence interval for mean (exponential distribution)

```{r}
library(tidyverse)
library(statBasics)
media_pop <- 800
taxa_pop <- 1 / media_pop
x <- rexp(100, rate = taxa_pop)
ci_exp(x)
```

## Confidence interval for mean

Confidence interval using *t* distribution is suitable, even the distribution is not normal, as illustrated in the example bellow.

```{r}
library(tidyverse)
library(statBasics)
media_pop <- 50
x <- rpois(100, lambda  = media_pop)
ci_general(x)
```
